---
title: "relational_data"
author: "ashley"
date: "November 2, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(nycflights13)
library(Lahman)
library(babynames)
library(nasaweather)
library(fueleconomy)
library(ggplot2)
library(maps)
```

#13.2.1
1. In flights you will need the origin and dest, in the airports you will need lat and long. We would have to combine the flights table to the airports table for both the origin and destination so that you get the lat and long of both. 
2. The relationship between weather and airports is that origin and faa is the same thing.
3. The same variables would match together. So year, month, day, hour, and origin would match. 
4. Add a table in flights that show the special ones with the day year and month so that it symbolizes the day of special flights. 

```{r flights}
airports
airlines
planes
flights
weather
```

#13.3.1
1.
```{r flights_key}
flights %>% 
  count(year, month, day, tailnum) %>% 
  filter(n > 1)
```

2.batting: playerId, yearId,and stint
  babynames: year, sex, name
  atmos:lat, long, year, month
  vehicles: id
  diamonds:there is not a primary key
```{r keys}
Lahman::Batting%>%
  group_by(playerID, yearID, stint)

babynames::babynames%>%
  group_by(year, sex, name) 

nasaweather::atmos%>%
  group_by(lat, long, year, month)

fueleconomy::vehicles%>%
  group_by(id)

diamonds
```

3.

#13.4.6
1.
```{r map}
airports %>%
  semi_join(flights, c("faa" = "dest")) %>%
  ggplot(aes(lon, lat)) +
    borders("state") +
    geom_point(aes(color = dst)) +
    coord_quickmap()
```

2.
```{r combine}
flights%>%
  inner_join(airports, c("origin" = "faa"))%>%
  select(flight, origin, dest, lat, lon)
```
3. There is not an obvious relationship
```{r age}
flights%>%
    inner_join(airports, c("origin" = "faa"))%>%
  select(dep_delay, arr_delay, year)
```
4. wind and visibility make differences in the flight delay time
```{r weather}
#atmos
#flights
flights%>%
  inner_join(weather)%>%
  select(dep_delay, wind_speed, precip, pressure,visib)%>%
  arrange(dep_delay)
```
5.Google shows there were two tornados that delayed the flights. the data show there is high wind speeds and a lot of pressure- inferring rough weather causing delays.
```{r delays}
flights%>%
  inner_join(weather, c("year", "month" , "day" , "hour" , "origin"))%>%
  select(day,month, year, dep_delay, wind_speed, precip, pressure, visib)%>%
  filter(month==6)
```


#13.5.1
1. Airlines MQ and AA have by far the most flights are without a tailnum. The airlines are not matched with a tailnum. 
```{r no_tailnum}
flights %>%
  anti_join(planes, by = "tailnum") %>%
  count(carrier, sort = TRUE)
```
2.
```{r count}
planes%>%  
  inner_join(flights)%>%
  group_by(tailnum)%>%
  count()%>%
  filter(n>100)
```
3.
```{r vehicles}
fueleconomy::vehicles%>%
  semi_join(fueleconomy::common, c("make", "model"))
```
4. Most of the delayed flights are during the months of decemeber and november. So the winter weather has an effect on the flights being delayed. 
```{r 4}
flights%>%
  inner_join(weather)%>%
  group_by(day,month,year)%>%
  select(dep_delay)%>%
  count()%>%
  arrange(desc(n))
```

5. the first one looks at the flights with the destination that does not match with an origin. and the second one looks at the flights that does has an arrival but does not have a match with the destination or the flights that does not have a destination in the data. 
```{r 1}
anti_join(flights, airports, by = c("dest" = "faa"))
```
```{r 2}
anti_join(airports, flights, by = c("faa" = "dest"))
```
6. It makes sence that they have a relationship because the airlines control where the planes go and so they will have the data of those planes and flights. 
```{r 6}
flights%>%
  group_by(tailnum, carrier)%>%
  count()%>%
  filter(n()>1)%>%
  select(tailnum)
```

